Weight-adjusted discontinuous Galerkin methods: matrix-valued weights and linear elastic wave propagation

### 2016-12-02 docs/elasticity_notes/main.tex
adding more elasticity paper updates

\renewcommand{\Oh}{{\Omega_h}}
\renewcommand{\L}{L^2\LRp{\Omega}}
\newcommand{\Lk}{L^2\LRp{D^k}}
\newcommand{\Ldk}{L^2\LRp{\partial D^k}}
\newcommand{\Dhat}{\widehat{D}}
\newcommand{\Lhat}{L^2\LRp{\Dhat}}

@@

\newcommand{\note}[1]{{\color{blue}#1}}
\newcommand{\remark}[1]{\textbf{\color{red}#1}}

@@

\date{}
\author{Jesse Chan}
\title{Weight-adjusted discontinuous Galerkin methods: matrix-valued weights with application to elastic wave propagation}

@@

\maketitle
%\tableofcontents

@@

@@

SEM \cite{komatitsch2010high,wilcox2010high}

High order DG \cite{kaser2006arbitrary, dumbser2006arbitrary, de2007arbitrary,ye2016discontinuous}

Interior penalty formulations.  Analysis: Riviere et al.\ \cite{riviere2003discontinuous,riviere2007discontinuous} and Delcourte et al.\ \cite{delcourte2009high,delcourte2015analysis}, dispersion analysis \cite{de2008interior,antonietti2016high}.

GPU-accelerated DG methods \cite{klockner2009nodal,godel2010gpu,godel2010scalability}.  Also elastic wave propagation and seismic wave applications \cite{modave2015nodal,modave2016gpu}

High order approximations of material data \cite{castro2010seismic,mercerat2015nodal}, but not low storage or not energy stable.

@@

We begin with the linear elastic wave equation under small deformations in a domain $\Omega$.  These equations can be written a first order velocity-stress  system 
\begin{align*}
\rho \pd{\bm{v}}{t} &= \Div{\bm{S}}\\
\pd{\bm{S}}{t} &= \frac{1}{2} \bar{\bm{C}}\LRp{\Grad\bm{v} + \Grad\bm{v}^T},
\end{align*}
where $\bm{v}$ is the velocity, $\bm{S}$ is the stress tensor, $\rho$ is density, and $\bar{\bm{C}}$ is the symmetric constitutive stiffness tensor relating stress and strain.  These equations can also be rewritten as a symmetric hyperbolic system of PDEs \cite{hughes1978classical}  
\begin{align}
\rho \pd{\bm{v}}{t} &= \sum_{i=1}^d \bm{A}_i^T \pd{\bm{\sigma}}{\bm{x}_i}\nonumber \\
\bm{C}^{-1} \pd{\bm{\sigma}}{t} &= \sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i}.
\label{eq:symelas}
\end{align}
where $\bm{C}$ is the symmetric matrix form of the constitutive tensor $\bar{\bm{C}}$ and $\sigma$ is a vector of length $N_d = \frac{d(d+1)}{2}$, the number of unique entries of the stress tensor $\bm{S}$.  We note that the matrices $\bm{A}_i$ are constant, while $\rho$ and $\bm{C}$ can vary spatially.  Furthermore, we will assume that $\rho$ and $\bm{C}$ are positive-definite and bounded such that 
\begin{align*}
\rho_{\min} &\leq \rho(\bm{x}) \leq \rho_{\max} \\
0 < c_{\min} \leq \bm{x}^T\bm{C}\bm{x} &\leq c_{\max} < \infty, \qquad \bm{x} \in \mathbb{R}^{N_d},
\end{align*}
which also implies that $\bm{C}^{-1}$ is positive-definite and bounded.

@@

\section{An energy stable discontinuous Galerkin variational formulation}

The advantage of the symmetric first order formulation of the elastic wave equations (\ref{eq:symelas}) is that it is straightforward to derive an energy stable discontinuous Galerkin formulation.  We assume that the domain $\Omega$ is Lipschitz and exactly triangulated by a mesh $\Omega_h$, which consists of elements $D^k$.  We further assume that each element $D^k$ is the image of a reference element $\widehat{D}$ under the local elemental mapping 
\[
\bm{x}^k = \bm{\Phi}^k \widehat{\bm{x}},
\]
where $\bm{x}^k = \LRc{x^k,y^k,z^k}$ denote physical coordinates on $D^k$ and $\widehat{\bm{x}} = \LRc{\widehat{x},\widehat{y},\widehat{z}}$ denote coordinates on the reference element.  We denote the determinant of the Jacobian of $\bm{\Phi}^k$ as $J^k$.

We will approximate components over each element $D^k$ from an approximation space $V_h\LRp{D^k}$, which we defined as
\[
V_h\LRp{D^k} = \bm{\Phi}^k \circ V_h\LRp{\widehat{D}}.
\]
where $V_h\LRp{\widehat{D}}$ is an approximation space defined on the reference element.  For the remainder of this work, we will take $V_h\LRp{\widehat{D}} = P^N\LRp{\widehat{D}}$, where $P^N(\LRp{\widehat{D}})$ is the polynomial space of total degree $N$ on the reference simplex.  In two dimensions, 
\[
P^N\LRp{\widehat{D}} = \LRc{ \widehat{x}^i \widehat{y}^j, \quad 0 \leq i + j \leq N},
\] 
while in three dimensions, 
\[
P^N\LRp{\widehat{D}} = \LRc{ \widehat{x}^i \widehat{y}^j \widehat{z}^k, \quad 0 \leq i + j +k \leq N}.  
\]

We denote the $L^2$ inner product and norm over $D^k$ by $\LRp{\cdot,\cdot}_{L^2\LRp{D^k}}$, such that
\[
\LRp{\bm{f},\bm{g}}_{\Lk} = \int_{D^k} \bm{f}\cdot\bm{g} \diff x = \int_{\widehat{D}} \bm{f}\cdot\bm{g} J^k, \qquad \nor{\bm{f}}_{\Lk}^2 = \LRp{\bm{f},\bm{f}}_{\Lk},
\]
where $\bm{f},\bm{g}$ are vector-valued functions.  Global $L^2$ inner products and norms are using local $L^2$ inner products and norms
\[
\LRp{\bm{f},\bm{g}}_{\Oh} = \sum_{D^k\in \Oh} \int_{D^k} \LRp{\bm{f},\bm{g}}_{\Lk}, \qquad \nor{\bm{f}}_{\Oh}^2 = \sum_{D^k\in \Oh} \nor{\bm{f}}^2_{\Lk}.
\]

@@

We can now specify a DG formulation for the linear elastic wave equation (\ref{eq:symelas}).  Symmetric hyperbolic systems readily admit a DG formulation based on penalty fluxes, as detailed in \cite{chan2016short}.  For the linear elastic wave equation in symmetric first order form, this formulation is given as
\begin{align}
\sum_{D^k\in \Oh}\LRp{\rho \pd{\bm{v}}{t},\bm{w}}_{\Lk} &= \sum_{D^k\in \Oh}\LRp{\LRp{ \sum_{i=1}^d \bm{A}_i^T\pd{\bm{\sigma}}{\bm{x}_i},\bm{w}}_{\Lk}  + \LRa{\frac{1}{2}\bm{A}_n^T\jump{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{w}}_{\Ldk}} \nonumber\\
\sum_{D^k\in \Oh}\LRp{\bm{C}^{-1} \pd{\bm{\sigma}}{t},\bm{q}}_{\Lk} &= \sum_{D^k\in \Oh}\LRp{\LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{q}}_{\Ldk}},
\label{eq:dgform}
\end{align}
where $\tau_v,\tau_\sigma \geq 0$ are penalty parameters and $\bm{A}_n$ is the normal matrix defined on a face $f$ as $\bm{A}_n = \sum_{i=1}^d \bm{n}_i \bm{A}_i$.  In two dimensions, $\bm{A}_n$ is 
\[
\bm{A}_n =  \LRp{\begin{array}{cc}
{n}_x & 0\\
 0 & {n}_y\\
{n}_y & {n}_x
\end{array}}.
\]
while in three dimensions, $\bm{A}_n$ is
\[
\bm{A}_n =  \LRp{\begin{array}{ccc}
{n}_x & 0 & 0\\
 0 & {n}_y & 0\\
0&  0 & {n}_z \\
0 &{n}_z & {n}_y\\
{n}_z & 0 & {n}_x\\
{n}_y & {n}_z & 0\\
\end{array}}.  
\]

@@

For a face which lies on a boundary, velocity boundary conditions are imposed by setting 
\begin{align*}
\jump{\bm{v}} &= 2\LRp{\bm{v}_{\rm bc} - \bm{v}^- }, \qquad \jump{\bm{A}_n^T\bm{\sigma}}=\jump{\bm{S}\bm{n}} = 0, 
\end{align*}
while traction boundary conditions are imposed by
\begin{align*}
\jump{\bm{A}_n^T\bm{\sigma}} = \jump{\bm{S}\bm{n}} = 2\LRp{\bm{t}_{\rm bc} - \bm{S}^-\bm{n} }= 2\LRp{\bm{t}_{\rm bc} - \bm{A}_n^T\bm{\sigma}^- }, \qquad \jump{\bm{v}} = 0.
\end{align*}
The numerical fluxes are then computed using these modified jumps.

For problems which involve the truncation of infinite or large domains, absorbing boundary conditions are required.  For such cases, we impose simple non-reflective boundary conditions \cite{leveque2002finite} through jumps 
\begin{align*}
\jump{\bm{A}_n^T\bm{\sigma}} = \jump{\bm{S}\bm{n}} = - \bm{S}^-\bm{n} = -\bm{A}_n^T\bm{\sigma}^- , \qquad \jump{\bm{v}} = - \bm{v}^-. 
\end{align*}
We note that more accurate absorbing conditions can be imposed using, for example, perfectly matched layers \cite{berenger1994perfectly} or high order absorbing boundary conditions \cite{hagstrom2004new,modave2016high}.

@@

It is straightforward to show that the formulation (\ref{eq:dgform}) is energy stable for zero velocity and traction boundary conditions, as well as non-reflective boundary conditions.  Integrating by parts the velocity equations yields
\LRp{\rho \pd{\bm{v}}{t},\bm{w}}_{\Lk} &= -\LRp{ \sum_{i=1}^d \bm{\sigma},\bm{A}_i \pd{\bm{w}}{\bm{x}_i}}_{\Lk}  + \LRa{\bm{A}_n^T\avg{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{w}} \nonumber\\
\LRp{\bm{C}^{-1} \pd{\bm{\sigma}}{t},\bm{q}}_{\Lk} &= \LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{q}} \label{eq:dgform}

@@

Let $\Gamma_h$ denote the set of unique faces in $\Oh$, and let $\Gamma_v,\Gamma_\sigma,\Gamma_{\rm abc}$ denote the parts of the boundary where velocity, traction, and non-reflective boundary conditions are imposed, respectively.  We separate surface terms into contributions from interior shared faces and from boundary faces.  On interior shared faces, 

@@

%\label{eq:stability}

@@

Combining these face contributions yields

@@

\label{eq:stability}

@@

Since $\rho$ and $\bm{C}^{-1}$ are positive definite, the left hand side of (\ref{eq:stability}) is an $L^2$-equivalent norm on $(\bm{u},\bm{\sigma})$ 
\[
C_1 \nor{(\bm{u},\bm{\sigma})}_{\L}^2 \leq \sum_{D^k\in \Oh}\frac{1}{2}\pd{}{t}&\LRp{\LRp{\rho \bm{v},\bm{v}}_{\Lk} + \LRp{\bm{C}^{-1} \bm{\sigma},\bm{\sigma}}_{\Lk}} \leq C_2 \nor{(\bm{u},\bm{\sigma})}_{\L}^2
\]
for $\nor{(\bm{u},\bm{\sigma})}_{\L}^2 = \nor{\bm{u}}^2_{\L} + \nor{\bm{\sigma}}^2_{\L}$.  
% $\nor{(\bm{u},\bm{\sigma})}_{\L}$ is 
%\[
%\nor{(\bm{u},\bm{\sigma})}_{\L}^2 =\nor{(\bm{u},\bm{\sigma})}_{\L}^2 = \nor{\bm{u}}^2_{\L} + \nor{\bm{\sigma}}^2_{\L}.
%\]
This implies that the DG formulation is energy stable and that the magnitude of the DG solution $(\bm{u},\bm{\sigma})$ is non-increasing in time.  This also shows that dissipation present for positive penalization constants $\tau_v,\tau_{\sigma}$ acts on non-conforming components with non-zero jumps $\bm{A}_n^T\jump{\bm{v}}$ and $\bm{A}_n\jump{\bm{\sigma}}$.  In fact, it was shown in \cite{chan2016short} that, in the limit as $\tau_v,\tau_\sigma \rightarrow\infty$, the eigenspaces of DG discretizations split into a conforming part consisting of $\bm{u},\bm{\sigma}$ which satisfy
\[
\bm{A}_n\jump{\bm{u}} = 0, \qquad \bm{A}_n^T\jump{\bm{\sigma}} = 0
\]
and an non-conforming part (defined through the $L^2$ orthogonal complement) corresponding to eigenvalues contain real parts which approach $-\infty$.  For the linear elastic wave equations, these conditions are equivalent to the requirement of $C^0$ continuity for $\bm{u}$ and normal continuity of the stress tensor $\jump{\bm{S}}\bm{n} = 0$.
%\begin{align*}
%\jump{\bm{S} \bm{n}} = 0.
%\jump{\bm{\sigma}_{xx} \bm{n}_x + \bm{\sigma}_{xy} \bm{n}_y + \bm{\sigma}_{xz} \bm{n}_z} &= 0\\
%\jump{\bm{\sigma}_{xy} \bm{n}_x + \bm{\sigma}_{yy} \bm{n}_y + \bm{\sigma}_{yz} \bm{n}_z} &= 0\\
%\jump{\bm{\sigma}_{xz} \bm{n}_x + \bm{\sigma}_{yz} \bm{n}_y + \bm{\sigma}_{zz} \bm{n}_z} &= 0.
%\end{align*}

While this DG formulation is energy stable, 
%Symmetric hyperbolic system
%\[
%\bm{A}_0\pd{\bm{v}}{t} = \sum_{i=1}^d \bm{A}_i\pd{\bm{v}}{\bm{x}_i}.
%\]
%The matrix $\bm{A}_0$ is given as
%\[
%\bm{A}_0 = \LRp{
%\begin{array}{cc}
%\rho\bm{I} &\\
%& \bm{C}^{-1}
%\end{array}
%}.
%\]

>>>>

\renewcommand{\Oh}{{\Omega_h}}
\renewcommand{\L}{L^2\LRp{\Omega}}
%\newcommand{\Lk}{\Lk}
%\newcommand{\Ldk}{L^2\LRp{\partial D^k}}
\newcommand{\Lk}{D^k}
\newcommand{\Ldk}{\partial D^k}
\newcommand{\Dhat}{\widehat{D}}
\newcommand{\Lhat}{L^2\LRp{\Dhat}}

@@

\newcommand{\note}[1]{{\color{blue}{#1}}}
\newcommand{\remark}[1]{\textbf{\color{red}#1}}

@@

\date{}
\author{Jesse Chan}
\title{Weight-adjusted discontinuous Galerkin methods: matrix-valued weights and linear elastic wave propagation}

@@

\maketitle

@@

%\tableofcontents

@@

SEM \cite{komatitsch2010high}

DG-SEM \cite{wilcox2010high}

High order DG \cite{kaser2006arbitrary, dumbser2006arbitrary, de2007arbitrary, ye2016discontinuous}

Interior penalty formulations.  Analysis: Riviere et al.\ \cite{riviere2003discontinuous,riviere2007discontinuous} and Delcourte et al.\ \cite{delcourte2009high,delcourte2015analysis}

GPU-accelerated DG methods \cite{klockner2009nodal,godel2010gpu,godel2010scalability}.  Also elastic wave propagation and seismic wave applications \cite{modave2015nodal, modave2016gpu}

High order approximations of material data \cite{castro2010seismic, mercerat2015nodal}, but not low storage or not energy stable.

@@

We begin with the linear elastic wave equation under small deformations in a domain $\Omega$.  These equations can be written a first order velocity-stress system for velocity $\bm{v}$ and symmetric stress tensor $\bm{S}$ 
\begin{align*}
\rho \pd{\bm{v}}{t} &= \Div{\bm{S}} + \bm{f}\\
\pd{\bm{S}}{t} &= \frac{1}{2} \bar{\bm{C}}\LRp{\Grad\bm{v} + \Grad\bm{v}^T},
\end{align*}
where $\bm{f}$ is the body force per unit volume, $\rho$ is density, and $\bar{\bm{C}}$ is the symmetric constitutive stiffness tensor relating stress and strain.  These equations can also be rewritten as a symmetric hyperbolic system of PDEs \cite{hughes1978classical}  
\begin{align}
\rho \pd{\bm{v}}{t} &= \sum_{i=1}^d \bm{A}_i^T \pd{\bm{\sigma}}{\bm{x}_i}\nonumber + \bm{f}\\
\bm{C}^{-1} \pd{\bm{\sigma}}{t} &= \sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i}.
\label{eq:symelas}
\end{align}
where $\bm{C}$ is the symmetric matrix form of the constitutive tensor $\bar{\bm{C}}$ and $\sigma$ is a vector of length $N_d = \frac{d(d+1)}{2}$, the number of unique entries of the stress tensor $\bm{S}$.  We note that the matrices $\bm{A}_i$ are constant, while $\rho$ and $\bm{C}$ can vary spatially.  Furthermore, we will assume that $\rho$ and $\bm{C}$ are positive-definite and bounded such that 
\begin{align*}
\rho_{\min} &\leq \rho(\bm{x}) \leq \rho_{\max} \\
0 < c_{\min} \leq \bm{x}^T\bm{C}\bm{x} &\leq c_{\max} < \infty, \qquad \bm{x} \in \mathbb{R}^{N_d},
\end{align*}
which also implies that $\bm{C}^{-1}$ is positive-definite and bounded.

@@

\section{An energy stable discontinuous Galerkin formulation for elastic wave propagation}

The advantage of the symmetric first order formulation of the elastic wave equations (\ref{eq:symelas}) is that it is straightforward to derive an energy stable discontinuous Galerkin formulation.  We assume that the domain $\Omega$ is Lipschitz and exactly triangulated by a mesh $\Omega_h$, which consists of elements $D^k$.  We further assume that each element $D^k$ is the image of a reference element $\widehat{D}$ under the local elemental mapping 
\[
\bm{x}^k = \bm{\Phi}^k \widehat{\bm{x}},
\]
where $\bm{x}^k = \LRc{x^k,y^k,z^k}$ denote physical coordinates on $D^k$ and $\widehat{\bm{x}} = \LRc{\widehat{x},\widehat{y},\widehat{z}}$ denote coordinates on the reference element.  We denote the determinant of the Jacobian of $\bm{\Phi}^k$ as $J$, and refer to it as the Jacobian for the remainder of this work.

We will approximate components over each element $D^k$ from an approximation space $V_h\LRp{D^k}$, which we define as the composition of the mapping $\bm{\Phi}^k$ and a reference approximation space $V_h\LRp{\widehat{D}}$
\[
V_h\LRp{D^k} = \bm{\Phi}^k \circ V_h\LRp{\widehat{D}}.
\]
For the remainder of this work, we will take $V_h\LRp{\widehat{D}} = P^N\LRp{\widehat{D}}$, where $P^N\LRp{\widehat{D}}$ is the polynomial space of total degree $N$ on the reference simplex.  In two dimensions, 
\[
P^N\LRp{\widehat{D}} = \LRc{ \widehat{x}^i \widehat{y}^j, \quad 0 \leq i + j \leq N},
\] 
and in three dimensions, 
\[
P^N\LRp{\widehat{D}} = \LRc{ \widehat{x}^i \widehat{y}^j \widehat{z}^k, \quad 0 \leq i + j +k \leq N}.  
\]

We denote the $L^2$ inner product and norm over $D^k$ by $\LRp{\cdot,\cdot}_{\Lk}$, such that
\[
\LRp{\bm{f},\bm{g}}_{\Lk} = \int_{D^k} \bm{f}\cdot\bm{g} \diff x = \int_{\widehat{D}} \bm{f}\cdot\bm{g} J, \qquad \nor{\bm{f}}_{\Lk}^2 = \LRp{\bm{f},\bm{f}}_{\Lk},
\]
where $\bm{f},\bm{g}$ are vector-valued functions.  Global $L^2$ inner products and norms are using local $L^2$ inner products and norms
\[
\LRp{\bm{f},\bm{g}}_{\Oh} = \sum_{D^k\in \Oh} \int_{D^k} \LRp{\bm{f},\bm{g}}_{\Lk}, \qquad \nor{\bm{f}}_{\Oh}^2 = \sum_{D^k\in \Oh} \nor{\bm{f}}^2_{\Lk}.
\]

@@

We can now specify a DG formulation for the linear elastic wave equation (\ref{eq:symelas}).  Symmetric hyperbolic systems readily admit a DG formulation based on penalty fluxes, as detailed in \cite{chan2016short}.  For the linear elastic wave equation in symmetric first order form, this formulation is given as
\begin{align}
\sum_{D^k\in \Oh} \LRp{\rho \pd{\bm{v}}{t},\bm{w}}_{\Lk} &= 
\sum_{D^k\in \Oh} \LRp{\LRp{ \sum_{i=1}^d \bm{A}_i^T\pd{\bm{\sigma}}{\bm{x}_i} + \bm{f},\bm{w}}_{\Lk}  + \LRa{\frac{1}{2}\bm{A}_n^T\jump{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{w}}_{\Ldk}} \nonumber\\
\sum_{D^k\in \Oh}\LRp{\bm{C}^{-1} \pd{\bm{\sigma}}{t},\bm{q}}_{\Lk} &=
\sum_{D^k\in \Oh}\LRp{\LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{q}}_{\Ldk}},
\label{eq:dgform}
\end{align}
for all $\bm{w},\bm{q}\in V_h\LRp{\Omega_h}$.  Here, $\tau_v,\tau_\sigma \geq 0$ are penalty parameters and $\bm{A}_n$ is the normal matrix defined on a face $f$ as $\bm{A}_n = \sum_{i=1}^d \bm{n}_i \bm{A}_i$.  In two dimensions, $\bm{A}_n$ is 
\[
\bm{A}_n =  \LRp{\begin{array}{cc}
{n}_x & 0\\
 0 & {n}_y\\
{n}_y & {n}_x
\end{array}}.
\]
while in three dimensions, $\bm{A}_n$ is
\[
\bm{A}_n =  \LRp{\begin{array}{ccc}
{n}_x & 0 & 0\\
 0 & {n}_y & 0\\
0&  0 & {n}_z \\
0 &{n}_z & {n}_y\\
{n}_z & 0 & {n}_x\\
{n}_y & {n}_z & 0\\
\end{array}}.  
\]

@@

For a face which lies on a boundary, velocity boundary conditions are imposed by setting 
\begin{align*}
\jump{\bm{v}} &= 2\LRp{\bm{v}_{\rm bc} - \bm{v}^- }, \qquad \jump{\bm{A}_n^T\bm{\sigma}}=\jump{\bm{S}\bm{n}} = 0, 
\end{align*}
while traction boundary conditions are imposed by
\begin{align*}
\jump{\bm{A}_n^T\bm{\sigma}} = \jump{\bm{S}\bm{n}} = 2\LRp{\bm{t}_{\rm bc} - \bm{S}^-\bm{n} }= 2\LRp{\bm{t}_{\rm bc} - \bm{A}_n^T\bm{\sigma}^- }, \qquad \jump{\bm{v}} = 0.
\end{align*}
For problems which involve the truncation of infinite or large domains, absorbing boundary conditions are required.  For such cases, we impose simple non-reflective boundary conditions \cite{leveque2002finite} through jumps 
\begin{align*}
\jump{\bm{A}_n^T\bm{\sigma}} = \jump{\bm{S}\bm{n}} = - \bm{S}^-\bm{n} = -\bm{A}_n^T\bm{\sigma}^- , \qquad \jump{\bm{v}} = - \bm{v}^-. 
\end{align*}
We note that more accurate absorbing conditions can be imposed using, for example, perfectly matched layers \cite{berenger1994perfectly} or high order absorbing boundary conditions \cite{hagstrom2004new,modave2016high}.

In all cases, boundary conditions are imposed by computing numerical fluxes using these modified jumps.  This imposition guarantees energy stability for all three sets of boundary conditions.

@@

It is straightforward to show that the formulation (\ref{eq:dgform}) is energy stable for zero velocity and traction boundary conditions, as well as non-reflective boundary conditions.  Integrating by parts the velocity equations gives
\LRp{\rho \pd{\bm{v}}{t},\bm{w}}_{\Lk} &= -\LRp{ \sum_{i=1}^d \bm{\sigma},\bm{A}_i \pd{\bm{w}}{\bm{x}_i}}_{\Lk}  + \LRa{\bm{A}_n^T\avg{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{w}}_{\Ldk} \nonumber\\
\LRp{\bm{C}^{-1} \pd{\bm{\sigma}}{t},\bm{q}}_{\Lk} &= \LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{q}}_{\Ldk} 

@@

Let $\Gamma_h$ denote the set of unique faces in $\Oh$, and let $\Gamma_v,\Gamma_\sigma,\Gamma_{\rm abc}$ denote the parts of the boundary where velocity, traction, and non-reflective boundary conditions are imposed, respectively.  We separate surface terms into contributions from interior shared faces and from boundary faces.  On an interior shared face, we sum contributions from the two adjacent elements to yield

@@

%\label{thm:stability}

@@

Combining all face contributions together gives the following result:
\begin{theorem}
The DG formulation (\ref{eq:dgform}) is energy stable for $\tau_v,\tau_\sigma \geq 0$, in the sense that

@@

\label{thm:stability}

@@

\end{theorem}

Since $\rho$ and $\bm{C}^{-1}$ are positive definite, the left hand side of (\ref{thm:stability}) is an $L^2$-equivalent norm on $(\bm{u},\bm{\sigma})$, and Theorem~\ref{thm:stability} implies that the magnitude of the DG solution $(\bm{u},\bm{\sigma})$ is non-increasing in time.  This also shows that dissipation present for positive penalization constants $\tau_v,\tau_{\sigma}$ acts on non-conforming components with non-zero jumps $\bm{A}_n^T\jump{\bm{v}}$ and $\bm{A}_n\jump{\bm{\sigma}}$.  In fact, it was shown in \cite{chan2016short} that, in the limit as $\tau_v,\tau_\sigma \rightarrow\infty$, the eigenspaces of DG discretizations split into a conforming part consisting of $\bm{u},\bm{\sigma}$ which satisfy
\[
\bm{A}_n\jump{\bm{u}} = 0, \qquad \bm{A}_n^T\jump{\bm{\sigma}} = 0
\]
and an non-conforming part (defined through the $L^2$ orthogonal complement) corresponding to eigenvalues contain real parts which approach $-\infty$.  For the linear elastic wave equations, these conditions are equivalent to the requirement of $C^0$ continuity for $\bm{u}$ and normal continuity of the stress tensor $\jump{\bm{S}}\bm{n} = 0$.
%\begin{align*}
%\jump{\bm{S} \bm{n}} = 0.
%\jump{\bm{\sigma}_{xx} \bm{n}_x + \bm{\sigma}_{xy} \bm{n}_y + \bm{\sigma}_{xz} \bm{n}_z} &= 0\\
%\jump{\bm{\sigma}_{xy} \bm{n}_x + \bm{\sigma}_{yy} \bm{n}_y + \bm{\sigma}_{yz} \bm{n}_z} &= 0\\
%\jump{\bm{\sigma}_{xz} \bm{n}_x + \bm{\sigma}_{yz} \bm{n}_y + \bm{\sigma}_{zz} \bm{n}_z} &= 0.
%\end{align*}

\subsection{The semi-discrete matrix system for DG}

The solution to (\ref{eq:dgform}) can be approximated by discretizing in space and using an explicit time integrator, which requires only evaluations of local contributions over $D^k$ to the DG formulation
\begin{align}
\LRp{\rho \pd{\bm{v}}{t},\bm{w}}_{\Lk} &= {\LRp{ \sum_{i=1}^d \bm{A}_i^T\pd{\bm{\sigma}}{\bm{x}_i},\bm{w}}_{\Lk}  + \LRa{\frac{1}{2}\bm{A}_n^T\jump{\bm{\sigma}} + \frac{\tau_v}{2}\bm{A}_n^T\bm{A}_n\jump{\bm{v}},\bm{w}}_{\Ldk}} \nonumber\\
\LRp{\bm{C}^{-1} \pd{\bm{\sigma}}{t},\bm{q}}_{\Lk} &= {\LRp{\sum_{i=1}^d \bm{A}_i \pd{\bm{v}}{\bm{x}_i},\bm{q}}_{\Lk} + \LRa{\frac{1}{2}\bm{A}_n\jump{\bm{v}} + \frac{\tau_{\sigma}}{2}\bm{A}_n\bm{A}_n^T\jump{\bm{\sigma}},\bm{q}}_{\Ldk}}. 
\label{eq:dgloc} 
\end{align}

Let $\LRc{\phi_i}_{i=1}^{N_p}$ be a basis for $P^N\LRp{\widehat{D}}$.  We define the reference mass matrix $\widehat{\bm{M}}$ and the physical mass matrix $\bm{M}$ for an element $D^k$ as
\[
\LRp{\widehat{\bm{M}}}_{ij} = \int_{\widehat{D}}\phi_j\phi_i \diff{\widehat{\bm{x}}}, \qquad \LRp{\bm{M}}_{ij} = \int_{D^k} \phi_j\phi_i \diff{\bm{x}} = \int_{\widehat{D}}\phi_j\phi_i J\diff{ \widehat{\bm{x}}}.
\]
For affine mappings, $J$ is constant and $\bm{M} = J\widehat{\bm{M}}$.  We also define weak differentiation matrices $\bm{S}_{k}$ and face mass matrix $\bm{M}_f$ such that
\[
\LRp{\bm{S}_{k}}_{ij} =  \int_{D^k} \pd{\phi_j}{\bm{x}_k}\phi_i \diff{\bm{x}}, \qquad \LRp{\bm{M}_{f}}_{ij} =  \int_{f} \phi_j \phi_i \diff{\bm{x}} = \int_{\widehat{f}} \phi_j \phi_i J^f\diff{\widehat{\bm{x}}},
\]
where $J^f$ is the Jacobian of the mapping from a reference face $\widehat{f}$ to $f$.  For affinely mapped simplices, $J^f$ is also constant and $\bm{M}_f = J^f \widehat{\bm{M}}_f$, where the definition of the reference face mass matrix $\widehat{\bm{M}}_f$ is analogous to the definition of the reference mass matrix $\widehat{\bm{M}}$.

Finally, we define weighted mass matrices.  Let $w(\bm{x})\in \mathbb{R}$ and $\bm{W}(\bm{x})\in \mathbb{R}^{m\times n}$.  Then, scalar and matrix-weighted mass matrices $\bm{M}_w$ and $\bm{M}_{\bm{W}}$ are defined through
\[
\LRp{\bm{M}_w}_{ij} = \int_{D^k} w(\bm{x}) \phi_j(\bm{x})  \phi_i(\bm{x}) \diff x, \qquad \bm{M}_{\bm{W}} = \LRp{\begin{array}{ccc}
\bm{M}_{\bm{W}_{1,1}}& \ldots & \bm{M}_{\bm{W}_{1,n}}\\
\vdots & \ddots & \vdots\\
\bm{M}_{\bm{W}_{m,1}} & \ldots & \bm{M}_{\bm{W}_{m,n}}\\
\end{array}},
\]
where $\bm{M}_{\bm{W}_{i,j}}$ is the scalar weighted mass matrix weighted by the $(i,j)$ entry of $\bm{W}$.
